### 2015-10-20 pde/pdetest1.tex
added a lot of PDE stuff

\documentclass{article}
\usepackage{amssymb,amsmath}
\usepackage{mathtools}
\usepackage{graphicx}
\topmargin=-0.3in
\textheight=9.2in
\textwidth=168mm
\oddsidemargin=-0.2in
\evensidemargin=-0.2in
\begin{document}

@@

\indent We will begin these notes in a reverse order; that is, we will start by solving a second order PDE, which is probably the hardest problem we will be presented with. Let's go ahead and look at the heat equation, which is a nice one to solve.\\

@@

\indent \textbf{\textit{Solution:}} We know that out solution will be in the form of $ u(x,t) = X(x)T(t) $. If we take this fact at face value, we can use the separation of variables method to solve this. First, we substitute in our solution into the PDE:

@@

%\sqrt{\lambda}L = \sin^{-1}{0}

@@

\newpage
\section{First Order Partial Differential Equations}
\hrule
\noindent\\
\noindent\textbf{\textit{Ex:}} Solve the following PDE
\[
\begin{cases*}
yu_{x} + xu_{y} = 0\\
u(0,y) = e^{-y^{2}}
\end{cases*}
\]
\indent \textbf{\textit{Solution:}} This is a nice PDE; there are two ways to solve it, but we will focus on the easier one. We can see that we have a coefficient on both $u_{x}$ and $u_{y}$. This tells us that when $u$ was differentiated, there was a constant $y$ and $x$. So now we can say that
\[
\frac{dy}{dx} = \frac{x}{y}
\]
\noindent This is a simple ODE to solve. We can use separation of variables:
\begin{align*}
xdx &= ydy\\
\int{x dx} &= \int{y dy}\\
\frac{1}{2}x^{2} &= \frac{1}{2}y^{2} + c\\
x^{2} - y^{2} &= c_{0}\\
y^{2} - x^{2} &= c_{1}
\end{align*}
\noindent You'll notice that we multiplied both sides by $-1$. This is for convenience. Now we can express the answer as
\[
u(x,y) = f(y^{2} - x^{2})
\]
\noindent And that will suffice.

\end{document}

>>>>

\documentclass{article}
\usepackage{amssymb,amsmath}
\usepackage{mathtools}
\usepackage{graphicx}
\topmargin=-0.3in
\textheight=9.2in
\textwidth=168mm
\oddsidemargin=-0.2in
\evensidemargin=-0.2in
\begin{document}
%\indent Let's look at another second order PDE: the wave equation. This equation will be a little harder to solve due to the fact that it contains two second order derivates. Let's take a look:\\
%\noindent\textbf{\textit{Ex:}} Solve the following PDE
%\[
%\begin{cases*}
%u_{tt} - u_{xx} = 0\\
%u(0,t)=u(L,t) = 0\\
%u(x,0) = f(x),\ u_{t}(x,0) = g(x)
%\end{cases*}
%\]
%\indent Let's take a second before we jump into solving this equation to analyze it. We have a homogeneous, second order PDE. We also have boundary conditions, much like with the heat equation. The initial conditions are a little strange, but we won't get into them until we discuss Fourier series.\\
%\noindent We will use separation of variables to solve this equation. With this technique, we know that we are looking for a solution in the form of 
%\[
%u(x,t) = \xi(x)\phi(t).
%\]
%\noindent So, we substitute into the original PDE, and we get as follows:
%\begin{gather*}
%u_{tt} - u_{xx} = 0 =\phi^{''}\xi - \xi^{''}\phi\\
%\frac{\phi^{''}}{\phi} = \frac{\xi^{''}}{\xi} = \lambda
%\end{gather*}
%\noindent Now we need to find our eigenvalues and eigenfunctions. There's a trick here, however. In the heat equation, we had the liberty of solving the time equation whenever we wanted to. Here, we don't have that ability just yet, since $\phi(t)$ is ``trapped'' in a second order ODE. The only real choice we have here is to do what we did earlier, and apply our boundary conditions to the $\xi(x)$ ODE. We won't delve into the process for finding the eigenvalues, so here's the only non-trivial pair:
%\[
%\{\lambda_{n} = (n\pi)^{2},\ \sin{(n\pi x)}\}
%\]
%\indent Now let's take a second to imagine that we had had boundary conditions for our second order ODE. We would have had the same setup for finding the eigenvalues, but our eigenfunction would have involved $t$ instead of $x$.

\newpage
\section{First Order Partial Differential Equations}
\hrule
\noindent\\
\noindent\textbf{\textit{Ex:}} Solve the following PDE
\[
\begin{cases*}
yu_{x} + xu_{y} = 0\\
u(0,y) = e^{-y^{2}}
\end{cases*}
\]
\indent \textbf{\textit{Solution:}} This is a nice PDE; there are two ways to solve it, but we will focus on the easier one. We can see that we have a coefficient on both $u_{x}$ and $u_{y}$. This tells us that when $u$ was differentiated, there was a constant $y$ and $x$. So now we can say that
\[
\frac{dy}{dx} = \frac{x}{y}
\]
\noindent This is a simple ODE to solve. We can use separation of variables:
\begin{align*}
xdx &= ydy\\
\int{x dx} &= \int{y dy}\\
\frac{1}{2}x^{2} &= \frac{1}{2}y^{2} + c\\
x^{2} - y^{2} &= c_{0}\\
y^{2} - x^{2} &= c_{1}
\end{align*}
\noindent You'll notice that we multiplied both sides by $-1$. This is for convenience. Now we can express the answer as
\[
u(x,y) = f(y^{2} - x^{2})
\]
\noindent And we are done.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{The Heat and Wave Equations}
\indent Let's take a look at the second order PDE. We will start with a homogeneous equation, which we will tackle with the separation of variables method. The general strategy is that we want to take our PDE and convert it to an ODE, which we know how to solve. The problems presented here also involve boundary conditions, which will require us to solve for eigenvalues/functions. We will also be dealing with series (specifically Fourier Series) to help us solve these problems. Let's start by looking at the homogeneous heat equation.\\

@@

\indent \textbf{\textit{Solution:}} We will use the separation of variable technique to solve this problem. We know to use separation of variable because we are dealing with a homogeneous PDE with homogenous boundary conditions. Separation of variables tells us that our solution will be in the form of $ u(x,t) = X(x)T(t) $. If we take this fact at face value, we can use the separation of variables method to solve this. First, we substitute in our solution into the PDE:

@@

@@

\newpage

\indent It will benefit us to take a moment and discuss how boundary conditions effect our Fourier series. We have three different combinations of boundary conditions, where each one will produce a different result in our series solution.
\begin{enumerate}
\item \textit{Dirichlet Boundary Conditions}\\
Dirichlet boundary conditions involve the function for which we are looking. In other words, the boundary conditions are presented in terms of the functions we are trying to find, rather than the derivative of that function. Typically, this type of boundary condition will take the form of (in the homogeneous case)
\[u(0,t) = u(L,t) = 0.\]
In regards to our Fourier series, we expect to see $\sin{(\frac{n\pi x}{L})}$. Our basis will look like
\[
\left\{ \lambda_{n} = \left(\frac{n\pi}{2}\right)^{2}, \phi_{n}(x) = \sin{\left(\frac{n\pi x}{L}\right)} \right\}
\]
\item \textit{Neumann Boundary Conditions}\\
Neumann boundary conditions involve the derivatives of the function for which we are solving. Neumann boundary conditions will look like
\[u_{x}(0,t) = u_{x}(L,t) = 0.\]
This will cause our Fourier series to contain $\cos{(\frac{n\pi x}{L})}$. Out basis will look like
\[
\left\{ \lambda_{n} = \left(\frac{n\pi}{2}\right)^{2}, \phi_{n}(x) = \cos{\left(\frac{n\pi x}{L}\right)} \right\}
\]
\item \textit{Robin Boundary Conditions}\\
Robin boundary conditions will contain a mixture of the function and it's derivative. Typically, we will see
\[ u_{x}(0,t) = u(L,t) = 0\qquad \text{or}\qquad u(0,t) = u_{x}(L,0) = 0.\]
These boundary conditions will give us two different results, depending on whether we have the derivative first or second. We will have
\begin{align*}
u_{x}(0,t) = u(L,t) = 0 &\implies \left\{\lambda_{n} = \left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}\right)^{2}, \phi_{n}(x) = \cos{\left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}x\right)}\right\}\\
u(0,t) = u_{x}(L,t) = 0 &\implies \left\{\lambda_{n} = \left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}\right)^{2}, \phi_{n}(x) =\sin{\left(\left[n + \frac{1}{2}\right]\frac{\pi}{L}x\right)}\right\}
\end{align*}
\end{enumerate}

\indent Now we can take a look at a slightly harder PDE. We still have the heat equation, but we will not have homogeneous boundary conditions. Our goal here is to find a way to transform our boundary conditions into a homogenous form and take it from there.\\\\
\begin{cases*}
u_{t} - u_{xx} = 0\\
u(0,t) = 0,\ u(1,t) = t\\
u(x,0) = x^{2}
\end{cases*}
\indent \textbf{\textit{Solution:}} We would like to just jump into this PDE and solve it, but it's not quite that simple. We have to find a way to express our boundary conditions in a homogeneous manner. We can do this through a clever substitution; we will create a new PDE, denoted $\Upsilon$. To find $\Upsilon$, we use the following formula:
\[ \Upsilon = B_{1} + \frac{x}{l}\Delta B, \]
\noindent where $B$ is a boundary condition. In our case, we have as follows:
\Upsilon &= B_{1} + \frac{x}{l}\Delta B\\
&= 0 + \frac{x}{1}(t - 0)\\
&= xt
\noindent Now, we find a new PDE, denoted by $\upsilon$. We have another formula that tells us
\[ \upsilon(x,t) = u(x,t) - \Upsilon(x,t) \]
\noindent Applying this formula gives us
\upsilon(x,t) &= u(x,t) - \Upsilon\\
&= u(x,t) - xt
\noindent Let's take a moment to analyze the result. We wanted to find a new PDE in terms of our old one that would allow us to have homogeneous boundary conditions. We found a solution, called $\upsilon$, in terms of our original PDE, called $u$. Our original PDE was $u_{t} - u_{xx} = 0$, so to properly perform our substitution, we need to find $\upsilon_{t}$ and $\upsilon_{xx}$. Through some careful differentiation, we have
\[ \upsilon_{t} = u_{t} - x \qquad\text{and}\qquad \upsilon_{xx} = u_{xx}\]
\noindent To find our new PDE, we simply evaluate.
\upsilon_{t} - \upsilon_{xx} = u_{t} - x - u_{xx} = u_{t} - u_{xx} - x = 0 - x = -x
\noindent So we finally have a new PDE that looks like
\begin{cases*}
\upsilon_{t} - \upsilon_{xx} = -x\\
\upsilon(0,t) = \upsilon(1,t) = 0\\
\upsilon(x,0) = x^{2}
\end{cases*}
\noindent You may wonder how this has made anything simpler, since the PDE itself is no longer homogeneous. By forcing the boundary conditions to be homogeneous, we can now use a Fourier series to find our solution. Recall that earlier we discussed the boundary condition and their resulting eigenvalue/eigenfunction. Here, we have Dirichlet boundary conditions, which tells us that we have a Fourier sine series. We have that
\upsilon(x,t) &= \sum_{n = 1}^{\infty}\left[a_{n}(t)\sin{(n\pi x)}\right]\\
\upsilon_{t}(x,t) &= \sum_{n = 1}^{\infty}\left[a_{n}'(t)\sin{(n\pi x)}\right]\\
\upsilon_{xx}(x,t) &= -\sum_{n = 1}^{\infty}\left[a_{n}(t)(n\pi)^{2}\sin{(n\pi x)}\right]\\ 
\noindent Which tells us that
\upsilon_{t} - \upsilon_{xx} = \sum_{n = 1}^{\infty}\left[(a_{n}'(t) + (n\pi)^{2}a_{n}(t))\sin{(n\pi x)}\right] = -x
\noindent This may look rather complicated, but all we have is a function's Fourier sine series. Essentially, we have
-x = \sum_{n = 1}^{\infty}\left[ b_{n}\sin{(n\pi x)} \right]
\noindent To find $b_{n}$, we use the formula that was discussed in the earlier section about Fourier series. We have that
a_{n}'(t) + (n\pi)^{2}a_{n}(t) &= \int_{0}^{1}-x\sin{(n\pi x)}dx\\
&= \frac{2x}{n\pi}\cos{(n\pi x)}\bigg\vert_{0}^{1}\\
&= \frac{2}{n\pi}
\noindent We now have an ODE which we can solve for $a_{n}$. Note that from out PDE, we do have an initial condition for the ODE, but it is not as simple as just saying $a_{n}(0) = x^{2}$. At $t = 0$, we have the following series:
\sum_{n=1}^{\infty} a_{n}(0)\sin{(n\pi x)}
\noindent We got this by taking our original $\upsilon(x,t)$ and plugging in $0$ for $t$. We know how to get $a_{n}(0)$, since this is just a simple Fourier sine series. It is given by
a_{n}(0) &= 2\int_{0}^{1}x^{2}\sin{(n\pi x)}dx\\
&= \frac{(-1)^{(n + 1)}}{n\pi} + \frac{2}{(n\pi)^{3}}\left[ (-1)^{n} - 1 \right]
\noindent We now have all we need to solve the ODE. We will omit the work for the solution, but the final answer for $a_{n}$ is 
a_{n}(t) = \frac{4\left[ (-1)^{n} - 1 \right]}{(n\pi)^{3}}e^{-(n\pi)^{2}t} + \frac{2(-1)^{n}}{(n\pi)^{3}}
\noindent Now all that's left to do is go back to $u$ from $\upsilon$. This may seem hard, but we saw earlier that $\upsilon = u - \Upsilon$. Solving for $u$ gives us $u = \upsilon + \Upsilon$. Recall that $\Upsilon$ was $xt$, and $\upsilon$ was the series for which we just found $a_{n}$. Putting it all together gives us
u(x,t) = xt + \sum_{n=1}^{\infty}\left[\frac{4\left[ (-1)^{n} - 1 \right]}{(n\pi)^{3}}e^{-(n\pi)^{2}t} + \frac{2(-1)^{n}}{(n\pi)^{3}}\right]\sin{(n\pi x)}
\noindent And we are done.
\newpage

\noindent \textbf{\textit{Ex:}} Solve the following PDE
\[
\begin{cases*}
u_{tt} - u_{xx} = 1 + t\cos{(2\pi x)}\\
u_{x}(0,t) = u_{x}(1,t) = 0\\
u(x,0) = 2\cos{(2\pi x)},\quad u_{t}(x,0) = 1 + \cos{(\pi x)} - \cos{(2\pi x)}
\end{cases*}
\]
\indent \textbf{\textit{Solution:}} This problem is an example of the wave equation with a non-homogeneous solution. We are lucky this time; the boundary conditions are already homogeneous, so there is no need to change our variable. We know that since both of the boundary conditions are derivatives, we are dealing with a Fourier cosine series. This yields the eigenvalue/function of
\[
\left\{\lambda_{n} = (n\pi)^{2}, \cos{(n\pi x)}\right\}
\]
Since we are dealing with a cosine series, we also have a leading term, $a_{0}$. Our full series looks like
\[
u(x,t) = a_{0}(t) + \sum_{n=1}^{\infty} a_{n}(t)\cos{(n\pi x)}
\]
We need $u_{tt}$ and $u_{xx}$, so we have:
\[
u_{tt}(x,t) = a_{0}''(t) + \sum_{n=1}^{\infty} a_{n}''(t)\cos{(n\pi x)} 
\qquad \text{and} \qquad 
u_{xx}(x,t) = -\sum_{n=1}^{\infty} (n\pi)^{2}a_{n}(t)\cos{(n\pi x)}
\]
All we have to do now is simply plug into our original equation. This gives us
\begin{align*}
a_{0}''(t) + \sum_{n=1}^{\infty} a_{n}''(t)\cos{(n\pi x)} + \sum_{n=1}^{\infty} (n\pi)^{2}a_{n}(t)\cos{(n\pi x)} &= 1 + t\cos{(2\pi x)}\\
a_{0}''(t) + \sum_{n=1}^{\infty}(a_{n}''(t) + (n\pi)^{2}a_{n}(t))\cos{(n\pi x)} &= 1 + t\cos{(2\pi x)}
\end{align*}
Now we can play the matching game. We have a $constant + series\ term$ on the right side, and $constant + series$ on the left, so we can match up the constants. This gives us 
\[
a_{0}''(t) = 1
\]
Now we match up the series term. On the right hand side, we have $\cos{(2\pi x)}$, so we know that for $n= 2$, we have
\[
a_{2}''(t) + 4\pi^{2}a_{2}(t) = t
\]
$a_{0}$ and $a_{2}$ are the only terms that appear, so we can assume for now that $\forall n \neq 2\ \text{and}\ 3, a_{n} = 0$. We are left with two ODE's, so we now need to find the corresponding initial conditions. Let's apply the first initial condition to our PDE. This gives us
\[a_{0}(0) + \sum_{n=1}^{\infty}a_{n}(0)\cos{(n\pi x)} = 2\cos{(2\pi t)}\]
Which tells us
\[a_{0} = 0 \qquad\qquad \qquad\qquad a_{2} = 2\]
Now we can apply the second initial condition. After taking the partial derivative of $u$ with respect to $t$, we have
\[
a_{0}'(0) + \sum_{n=1}^{\infty}a_{n}'(0)\cos{(n\pi x)} = 1 + \cos{(\pi x)} - \cos{(2\pi x)}
\]
Some more clever pattern matching give us
\[
a_{0}'(0) = 1 \qquad\qquad a_{1}'(0) = 1 \qquad\qquad a_{2}'(0) = -1
\]
Notice that we have a term $a_{1}'$. We cannot simply ignore this value; rather we have to take it into account when we find our solution. We know that we have an ODE for $a_{1}(t)$, but it may not be evident how to express it. Recall that earlier we said $\forall n \neq 2\ \text{and}\ 3, a_{n} = 0$. This means that $a_{1} = 0$, and since $a_{1}$ is a coefficient in the Fourier series, we know that $a_{1}''(t) + \pi^{2}a_{1}(t) = 0$. We now have three different ODE's that we must solve:
\begin{alignat*}{3}
&\begin{cases}
a_{0}''(t) = 1\\
a_{0}'(0) = 1,\quad a_{0}(0) = 1
\end{cases}
&\begin{cases}
a_{1}''(t) + \pi^{2}a_{1}(t) = 0\\
a_{1}'(0) = 1,\quad a_{1}(0) = 0
\end{cases}
&&\begin{cases}
a_{2}''(t) + 4\pi^{2}a_{2}(t) = t\\
a_{2}'(0) = 1,\quad a_{2}(0) = 2
\end{cases}
\end{alignat*}
Luckily, Mathematica is here to save the day. We will skip the work and let the computer do the thinking for us here. We get that
\begin{align*}
a_{0}(t) &= \frac{1}{2}t^{2} + t\\
a_{1}(t) &= \frac{\pi\cos{(t\pi)} + \sin{(t\pi)}}{\pi}\\
a_{2}(t) &= \frac{2\pi t + 16\pi^{3}\cos{(2\pi t)} - \sin{(2\pi t)} + 4\pi^{2}\sin{(2\pi t)}}{8\pi^{3}}
\end{align*}
Now all we need to do is plug back in to our $u$. Our final answer becomes
\[
u(x,t) = \frac{1}{2}t^{2} + t + \frac{\pi\cos{(t\pi)} + \sin{(t\pi)}}{\pi}\cos{(\pi x)} +  \frac{2\pi t + 16\pi^{3}\cos{(2\pi t)} - \sin{(2\pi t)} + 4\pi^{2}\sin{(2\pi t)}}{8\pi^{3}}\cos{(2\pi x)}
\]
And we are done.
\newpage

\subsection{The Laplace Equation}
\noindent \\
\indent Up to this point, we have only looked at the wave equation and the heat equation. There is, however, one other equation in the classical trinity: the Laplace Equation. We haven't dealt with this equation yet because we have to be very careful in our approach to solving it. When we looked at the homogeneous wave and heat equations, we used the general approach called \textit{Separation of Variables}. In the case of the homogeneous Laplace equation, we may not be able to tackle the problem using separation of variables. The problem arises from the various types of boundaries that effect the domain. We can either have a 'rectangular' bound, which will be separable, or a 'circular' bound which is \textbf{not} separable. Imagine that we have the following two problems:
\begin{align*}
&A =\begin{cases}
u_{xx} + u_{yy} = 0\quad \text{on}\ [0,1]\times[0,2]\\
u(x,0) = u(x,2) = 0\\
u(0,y) = y, u(1,y) = 0
\end{cases}
&B =\begin{cases}
u_{xx} + u_{yy} = 0\quad \text{on}\ x^{2} + y^{2} < a^{2}\\
u(x,y)\ \big |_{x^{2} + y^{2} = a^{2}} = f(x,y)
\end{cases}
\end{align*}
$A$ is an example of a separable Laplace equation. We can see graphically that the domain $[0,1] \times [0,2]$ is rectangular. $B$'s domain is not separable, since it is circular. Let's take a look $A$.\\\\
\noindent \textbf{\textit{Ex:}} Solve the following PDE
\[
\begin{cases*}
u_{xx} + u_{yy} = 0\quad \text{on}\ [0,1]\times[0,2]\\
u(x,0) = u(x,2) = 0\\
u(0,y) = y, u(1,y) = 0
\end{cases*}
\]
\indent \textbf{\textit{Solution:}} Earlier, we discussed the different boundary conditions we can encounter and their corresponding eigenvalues/functions. In this problem, we have Dirichlet boundary conditions. This tells us that
\[
\left\{\lambda_{n} = \left(\frac{n\pi}{2}\right)^{2}, \phi_{n}(y) = \sin{\left(\frac{n\pi y}{2}\right)}\right\}
\]
Notice that we used $y$ here instead of $x$. We did this because the boundary conditions are defined in terms of $y$, not $x$. Now we can continue to solve our problem. Recall that our separation of variables gives us two second order ODE's,
\[
X'' - \lambda X = 0 \qquad\qquad \text{and} \qquad\qquad \lambda Y - Y'' = 0
\]
We already solved for $Y(y)$ by finding the corresponding eigenvalues/vectors. Now we plug in our $\lambda_{n}$ into our $X$ ODE. Solving the resulting ODE will give us a basis of
\[
\left\{ e^{\frac{n\pi}{2}x}, e^{-\frac{n\pi}{2}x} \right\}
\]
Now we could continue at this point, but if we think ahead, we can see that when we eventually apply our initial conditions, nothing will cancel out. At this point, we want to transform our basis. We have as follows
\[
\left\{ e^{\frac{n\pi}{2}x}, e^{-\frac{n\pi}{2}x} \right\} \implies \left\{ \cosh{\left(\frac{n\pi x}{2}\right)}, \sinh{\left(\frac{n\pi x}{2}\right)} \right\}
\]
All we did here is apply an identity. Now we have everything we need to set up a solution. Our $u(x,y)$ will be
\[
u(x,y) = \sum_{n=1}^{\infty}\left[a_{n}\cosh{\left(\frac{n\pi x}{2}\right)} + b_{n}\sinh{\left(\frac{n\pi x}{2}\right)}\right]\sin{\left(\frac{n\pi y}{2}\right)}
\]
We can apply our initial conditions, which gives us $b_{n}\ \text{and}\ a_{n}$:
\begin{align*}
a_{n} &= \frac{4(-1)^{n+1}}{n\pi}\\
b_{n} &= -\coth{\left(\frac{n \pi}{2}\right)}\frac{4(-1)^{n+1}}{n\pi}
\end{align*}
This yields a final solution of
\[
u(x,y) = \sum_{n=1}^{\infty}\left[ \frac{4(-1)^{n+1}}{n\pi} \cosh{\left(\frac{n\pi x}{2}\right)} -\coth{\left(\frac{n \pi}{2}\right)}\frac{4(-1)^{n+1}}{n\pi}\sinh{\left(\frac{n\pi x}{2}\right)}\right]\sin{\left(\frac{n\pi y}{2}\right)}
\]
And we are done
\newpage

\indent Now we turn our attention to the Laplace Equation with a circular domain. As we discussed earlier, we can't simply say that our solution will be in the form of $X(x)Y(y)$. However, we can force the domain to be rectangular if we use polar coordinates. The translation is not too difficult. Let's consider $B$ from above.\\\\

\noindent \textbf{\textit{Ex:}} Solve the following PDE
\[
\begin{cases*}
u_{xx} + u_{yy} = 0\quad \text{on}\ x^{2} + y^{2} < a^{2}\\
u(x,y)\ \big |_{x^{2} + y^{2} = a^{2}} = f(x,y)
\end{cases*}
\]
\indent \textbf{\textit{Solution:}} We want to express the partial derivatives here in terms of $r\ \text{and}\ \theta$. Recall that we have
\[
x = r\cos{\theta} \qquad\qquad \text{and} \qquad\qquad y = r\sin{\theta}
\]
First, we will find the partial derivatives of $x$ and $y$ with respect to $r$ and $\theta$. Since $x = r\cos{\theta}$, 
\[
\frac{\partial x}{\partial r} = \cos{\theta} \qquad\qquad \frac{\partial x}{\partial \theta} = -r\sin{\theta}
\]
We can compute the derivatives for $y$ in a similar fashion:
\[
\frac{\partial y}{\partial r} = \sin{\theta} \qquad\qquad \frac{\partial y}{\partial \theta} = r\cos{\theta}
\]
Now we need to find $u_{rr}$ and $u_{\theta\theta}$. Let's focus on $u_{rr}$ first. We will need to start by finding $u_{r}$. We have as follows
\begin{align*}
\frac{\partial u}{\partial r}& = \frac{\partial u}{\partial x}\frac{\partial x}{\partial r} + \frac{\partial u}{\partial y}\frac{\partial y}{\partial r}\\
&= \cos{\theta}\frac{\partial u}{\partial x} + \sin{\theta}\frac{\partial u}{\partial y}
\end{align*}
We have $u_{r}$, so now we need to compute $u_{rr}$. We have as follows:
\begin{align*}
\frac{\partial^{2} u}{\partial r^{2}} &= \cos{\theta}\frac{\partial }{\partial r}\frac{\partial u}{\partial x} + \sin{\theta}\frac{\partial }{\partial r}\frac{\partial u}{\partial y}\\
&= \cos^{2}{\theta}\frac{\partial^{2} u}{\partial x^{2}} + 2\cos{\theta}\sin{\theta}\frac{\partial^{2}u}{\partial x\partial y} + \sin^{2}{\theta}\frac{\partial^{2} u}{\partial y^{2}}
\end{align*}
We won't delve into finding $u_{\theta\theta}$, as the partials get a little out of hand. However, we end up with
\begin{align*}
\frac{\partial^{2}u}{\partial r^{2}} + \frac{1}{r^{2}}\frac{\partial^{2}u}{\partial\theta^{2}} &= -\frac{1}{r}\frac{\partial u}{\partial r} + \frac{\partial^{2} u}{\partial x^{2}} + \frac{\partial^{2} u}{\partial y^{2}}\\
u_{xx} + u_{yy} &= u_{rr} + \frac{1}{r}u_{r} + \frac{1}{r^{2}}u_{\theta\theta} = 0
\end{align*}
We now have the Laplace equation represented in the polar coordinate system. Note that the translation of the boundary condition is simple. It becomes a function of only $\theta$. We have the following system:
\[
\begin{cases*}
u_{rr} + \frac{1}{r}u_{r} + \frac{1}{r^{2}}u_{\theta\theta} = 0\quad \text{on}\ 0 < r < a\\
u(\theta) = f(\theta)\quad \text{on}\ 0 < \theta < 2\pi
\end{cases*}
\]
Now we can use separation of variables to solve the PDE. The steps are the same (although this time we will use $R(r)\ \text{and}\ T(\theta)$). We will end up with the following equivalence:
\[
\frac{r^{2}R'' + rR'}{R} = \frac{T''}{T} = \lambda
\]
Since we have a Dirichlet boundary conditions, we know that
\[
\lambda_{n} = \frac{n^{2}}{4}
\]
Plugging this into our equation for $R$ gives us
\[
r^{2}R'' + rR' - \frac{n^{2}}{4}R = 0 
\]
\end{document}
